---
title: "Visualisation of factors on ratings of a race."
output:
  pdf_document: default
  html_document:
    df_print: paged
  word_document: default
---

```{r}
library(dplyr)
library(plyr)
library(ggplot2)
library(gridExtra)

```

---
In this R program we use the dataset which contains data collected from the 
Ergast developer API for F1 races. and the excitement index calculated using the
positive and negative sentiments of the tweets that are collected from twitter

The tweets are extracted using twitter api.

The Stanford CoreNLP is used to get the sentiments for the extracted tweets.

postiveSentiment in a tweet is calculated as the ratio of number of positive 
sentences and total sentences in a tweet.

negativeSentiment in a tweet is calculated as the ratio of number of negative
sentiment sentences and total sentences in a tweet.

All the positiveSentiments and negativeSentiments for all tweets of a 
particular race are summed to give totalpos and totalneg for a particular race.

As not all extracted tweets lead to a positive or negative sentiment we keep 
track of number of tweets which result in a positive or negative sentiment as
number of useful comments.

And finally the "Excitement index" is calculated as the ratio between 
(totalpos+totalneg) and (number of useful comments).

We use this dataset to visualize the effects of 3 variables namely 
number of overtakes(novertakes), proportion of unfinished drivers
(unfin/totaldrivers) - the drivers who started the race but could not finish due
to various reasons/total number of drivers and the weather of the 
race(wetweather) - F1 races are declared wet weather races if the weather 
conditions are wet/rainy/damp on the calculated "Excitement index".
---

```{r}
datax = read.csv("../Data/twitter_sentiments.csv")
colnames(datax)

```


```{r}
ggplot(data = datax,aes(x=Excitement_index,y= novertakes))+
  geom_point(aes(color=factor(wetweather)))+scale_color_manual(
    name= "Condition of track",
                                                          breaks = c(0,1),
                                                          values = c("orange","blue"),
                                                          labels= c("Normal","Wet"),
                                                          )+ylab("Number of overtakes in a race")+
  xlab("Excitement index of race")+
  geom_smooth(method = "lm", color = "red", se=FALSE)
```
```{r}
ggplot(data = datax,aes(x=Excitement_index,y= unfin,color=factor(wetweather)))+
  geom_point()+scale_color_manual(
    name= "Condition of track",
                                                          breaks = c(0,1),
                                                          values = c("orange","blue"),
                                                          labels= c("Normal","Wet"),
                                                          )+ylab("NUmber of unfinished drivers")+
  xlab("Excitement index of race")+geom_smooth(method = "lm",se=FALSE)
```


```{r}
ggplot(data = datax,aes(x=Excitement_index,y= unfin/totaldrivers))+
  geom_point(aes(color=factor(wetweather)))+scale_color_manual(
    name= "Condition of track",
                                                          breaks = c(0,1),
                                                          values = c("orange","blue"),
                                                          labels= c("Normal","Wet"),
                                                          )+ylab("Proportion of unfinished drivers")+
  xlab("Rating of race")+geom_smooth(method = "lm", color = "red", se=FALSE)
```


We perform a t-test to check if there is enough statistical significance to 
support the hypothesis that the fan ratings do not differ for wet weather 
and normal weather races.



```{r}
t.test(datax[datax$wetweather == 0,]$Excitement_index,datax[datax$wetweather == 1,]$Excitement_index,alternative = 'l')
```

Based on the results we can see that the null hypothesis which states that there 
no difference in fan ratings for wet weather and normal weather races can be 
rejected at a 0.05 significance level.


```{r}
datax$wetweather <- as.factor(datax$wetweather)
mu <- datax %>% group_by(wetweather) %>% summarise_at(vars(Excitement_index),list(name=mean))
mu
```
```{r}
ggplot(data=datax,aes(x=Excitement_index,fill=factor(wetweather)))+geom_density(alpha=0.5)+
  scale_fill_manual(
    name= "Condition of track",
    values = c("orange","blue"),labels=c("Normal","Wet")
  )+xlab("Excitement index of race")+
  ylab("Density")+
  geom_vline(data=mu,aes(xintercept=name,color = wetweather),linetype="dashed",show.legend=FALSE)
  #annotate(geom="text",label="p-value=",x=5,y=0.57)
```

```{r}
ggplot(data = datax,aes(y=Excitement_index,x= wetweather))+
  geom_boxplot(aes(group=wetweather,fill=factor(wetweather)))+scale_fill_manual(
    name= "Condition of track",
    values = c("orange","blue"),labels=c("Normal","Wet")
  )+xlab("Condition of Track")+
  ylab("Excitement index of race")
  #annotate(geom="text",label="p-value=0.0000001196",x=2.2,y=9.9)
```


```{r}
ggplot(data = datax,aes(sample=Excitement_index,color=factor(wetweather)))+
  stat_qq()+scale_color_manual(breaks = c(0,1),name= "Condition of track",
                                                          values = c("orange","blue"),
                                                          labels= c("Normal","Wet"),
                                                          )+
  geom_hline(yintercept = mean(datax[datax$wetweather==0,]$Excitement_index),color="orange")+
  geom_hline(yintercept = mean(datax[datax$wetweather==1,]$Excitement_index),color="blue")+
  ylab("Quantiles of race Excitement index")+xlab("Quantiles of normal distribution")

```
```{r}
dataxx <- datax %>% mutate(novertakes= (novertakes-mean(novertakes))/sd(novertakes))%>% 
  mutate(Excitement_index = (Excitement_index-mean(Excitement_index))/sd(Excitement_index)) %>% mutate(unfin = (unfin-mean(unfin))/sd(unfin))
lin=lm(Excitement_index ~ novertakes+unfin+wetweather,data = dataxx)
summary(lin)
```
```{r}

name = c("Intercept","Number of overtakes","Unfinished drivers","wet weather")
coef = lin$coefficients
conf1 = confint(lin)[,1]
conf2 = confint(lin)[,2]
df1 <- data.frame(name, coef, conf1, conf2)
df1
```
```{r}
ggplot(data=df1, aes(x=coef,y=name))+geom_errorbar(width = .1, aes(xmin= conf1, xmax=conf2))+
  geom_point(shape= 21, size =3, fill="blue")+
  geom_vline(xintercept = 0, linetype="dashed", color="red")+xlab("Coefficient of linear regression")+
  ylab("Regressors")
```

```{r}

# Create a dataframe only including the wet weather races
wet_weather_races <- datax[datax$wetweather==1,]

# Average total pos neg comments for wet weather races
avg_pos_neg_comments_wet <- (sum(wet_weather_races$Positive_tweets)+
                              sum(wet_weather_races$Negative_tweets))/
  sum(wet_weather_races$Useful_tweets)

# Average total pos neg comments for normal races
avg_pos_neg_comments_normal <-
  (sum(datax[datax$wetweather==0,]$Positive_tweets)+
     sum(datax[datax$wetweather==0,]$Negative_tweets))/
  sum(datax[datax$wetweather==0,]$Useful_tweets)

# Average excitement index for wet weather races
avg_excitement_index_wet <- sum(wet_weather_races$Excitement_index)/
  dim(wet_weather_races)[1]

# Average excitement index for normal races
avg_excitement_index_normal <- 
  sum(datax[datax$wetweather==0,]$Excitement_index)/
  dim(datax[datax$wetweather==0,])[1]

# Creating a dataframe with all the KPIs, so that they can be displayed in a
# plot.
comparison_wet_normal <- data.frame(c("Nass", "Normal", 
                                      "Nass", "Normal"))

names(comparison_wet_normal) <- "race_type"

comparison_wet_normal$values <- c(avg_pos_neg_comments_wet, 
                                  avg_pos_neg_comments_normal, 
                                  avg_excitement_index_wet, 
                                  avg_excitement_index_normal)

comparison_wet_normal$column <- c("pos_neg_comments", "pos_neg_comments", 
                                  "excitement_index", "excitement_index")

# Create a plot for average number of positive/negative comments
plot1 <- ggplot(data=comparison_wet_normal[comparison_wet_normal$column=="pos_neg_comments",], 
                aes(x=column, y=values, fill=race_type)) + 
  geom_bar(stat="identity", position=position_dodge()) + scale_fill_manual(values=c("#B5074A", "#432D57")) + theme(legend.position = "none", axis.text.x = element_blank(), axis.ticks.x = element_blank())  +
  xlab("Anteil positiver/negativer Tweets") + ylab("")

# Create a plot for the excitement_index
plot2 <- ggplot(data=comparison_wet_normal[comparison_wet_normal$column=="excitement_index",],
                aes(x=column, y=values, fill=race_type)) + 
  geom_bar(stat="identity", position=position_dodge()) + scale_fill_manual(values=c("#B5074A", "#432D57")) +
  theme(legend.position = "none", axis.text.x = element_blank(),
        axis.ticks.x = element_blank()) + xlab("Durchschnittlicher Excitement index") + ylab("")

# Create a legend for the plot
plot_legend <- ggplot(data=comparison_wet_normal[comparison_wet_normal$column=="pos_neg_comments",] 
                      ,aes(x=column, y=values, fill=race_type)) + 
  geom_bar(stat="identity",                                                 
           position=position_dodge()) + scale_fill_manual(values=c("#B5074A", "#432D57")) +
  theme(legend.position = "right", legend.direction = "horizontal", legend.title = element_blank())

# Some magic for creating a single legend for all the plots together
extract_legend <- function(my_ggp) {
  step1 <- ggplot_gtable(ggplot_build(my_ggp))
  step2 <- which(sapply(step1$grobs, function(x) x$name) == "guide-box")
  step3 <- step1$grobs[[step2]]
  return(step3)
}
shared_legend <- extract_legend(plot_legend)

# Create a grid with all the plots together and add the legend created above
grid.arrange(arrangeGrob(plot1,plot2, ncol = 2),
             shared_legend, nrow = 2, heights = c(10, 1))

```
