---
title: "Visualisation of factors on ratings of a race."
output:
  pdf_document: default
  html_document:
    df_print: paged
  word_document: default
---

```{r}
library(plyr)
library(dplyr)
library(ggplot2)
library(jpeg)
library(png)
library(grid)
```

---
In this R program we use the dataset which contains data collected from the 
Ergast developer API for F1 races. and the fan ratings collected from a kaggle
dataset. We use this dataset to visualize the effects of 3 variables namely 
number of overtakes(novertakes), number of unfinished drivers(unfin) - the 
drivers who started the race but could not finish due to various reasons and 
the weather of the race(wetweather) - F1 races are declared wet weather races if
the weather conditions are wet/rainy/damp.
---

```{r}
datax = read.csv("../Data/aggregated_basedOnFanRatings.csv")
colnames(datax)
```


```{r}

image=jpeg::readJPEG('Mercedes2.jpg')
ii=0.8
image <- matrix(rgb(image[,,1]*ii,image[,,2]*ii,image[,,3]*ii), nrow=dim(image)[1])

ggplot(data = datax)+
  theme_bw()+
  annotation_custom(rasterGrob(image = image,width = unit(1,"npc"), 
                               height = unit(1,"npc")), 
                               -Inf, Inf, -Inf, Inf)+ylab("Number of wins")+
  xlab("Year")+
  geom_bar(aes(x=factor(Year),fill=factor(constructorWon)),position = 
             'stack',stat = 'count',alpha=0.7)+
  scale_fill_manual(
    name= "Constructor",breaks = c(1,2,6,9,23,131),values = 
      c('orange','Grey','yellow2','red','olivedrab2','coral'),
      labels= c('McLaren','BMW Sauber','Ferrari','Red Bull','Brawn','Mercedes'))
  
  
```
```{r}
image=jpeg::readJPEG('Mercedes2.jpg')
ii=0.8
image <- matrix(rgb(image[,,1]*ii,image[,,2]*ii,image[,,3]*ii), nrow=dim(image)[1])

ggplot(data = datax)+
  theme_bw()+
  annotation_custom(rasterGrob(image = image,width = unit(1,"npc"), 
                               height = unit(1,"npc")), 
                               -Inf, Inf, -Inf, Inf)+ylab("Number of races")+
  xlab("Year")+
  geom_bar(aes(x=factor(Year),fill=factor(wetweather)),position = 
             'stack',stat = 'count',alpha=0.7)+
  scale_fill_manual(
    name= "Type of race",breaks = c(0,1),values = 
      c('orange','steelblue1'),
      labels= c('Normal weather','Wet weather'))


```




```{r}
image=png::readPNG('cup1.png')
ii=0.8
image <- matrix(rgb(image[,,1]*ii,image[,,2]*ii,image[,,3]*ii,image[,,4]), nrow=dim(image)[1])

databar = as.data.frame(table(datax$constructorWon))
databar$constructor <- c('McLaren','BMW Sauber','Ferrari','Red Bull','Brawn','Mercedes')

colors1 <- c('black','gold3','black','red','gold3','white')
colorss <- c('orange','Grey','yellow','white','white','grey27')
ggplot(data = databar,aes(x=reorder(Var1,Freq),y=Freq))+theme_bw()+
  annotation_custom(rasterGrob(image = image,width = unit(1,"npc"), 
                               height = unit(1,"npc")), 
                               -Inf, Inf, -Inf, Inf)+
  geom_bar(stat="identity",fill=colorss,color='black',alpha=0.8)+xlab("Constructors")+
  ylab("Number of wins")+
  geom_text(aes(label = constructor,y=1),fontface=2,color= colors1,angle=90,hjust=c(0.05,0,-0.8,-2,-0.4,-2))+
  theme_bw()+theme(axis.text.x = element_blank(),axis.text.y = element_blank(),
                               axis.ticks = element_blank())
```


```{r}
image=jpeg::readJPEG('rain1.jpg')
ii=0.5
image <- matrix(rgb(image[,,1]*ii,image[,,2]*ii,image[,,3]*ii), nrow=dim(image)[1])
ggplot(data = datax,aes(x=RATING,y= novertakes))+theme_bw()+
  annotation_custom(rasterGrob(image = image,width = unit(1,"npc"), 
                               height = unit(1,"npc")), 
                               -Inf, Inf, -Inf, Inf)+
  geom_point(aes(color=factor(wetweather)),alpha=3)+scale_color_manual(
    name= "Condition of track",
                                                          breaks = c(0,1),
                                                          values = c("orange","steelblue1"),
                                                          labels= c("Normal","Wet"),
                                                          )+ylab("Number of overtakes in a race")+
  xlab("Rating of race")+
  geom_smooth(method = "lm", color = "red", se=FALSE)
```
```{r}
image = jpeg::readJPEG("unfinished5.jpg")
ggplot(data = datax,aes(x=RATING,y= unfin))+theme_bw()+
  annotation_custom(rasterGrob(image = image,width = unit(1,"npc"), 
                               height = unit(1,"npc")), 
                               -Inf, Inf, -Inf, Inf)+
  geom_point(aes(color=factor(wetweather)),alpha=3)+scale_color_manual(
    name= "Condition of track",
                                                          breaks = c(0,1),
                                                          values = c("orange","steelblue1"),
                                                          labels= c("Normal","Wet"),
                                                          )+ylab("Number of unfinished drivers")+
  xlab("Rating of race")+geom_smooth(method = "lm", color = "red", se=FALSE)
```


```{r}
image = jpeg::readJPEG("unfinished5.jpg")
ggplot(data = datax,aes(x=RATING,y= unfin/totaldrivers))+theme_bw()+
  annotation_custom(rasterGrob(image = image,width = unit(1,"npc"), 
                               height = unit(1,"npc")), 
                               -Inf, Inf, -Inf, Inf)+
  geom_point(aes(color=factor(wetweather)))+scale_color_manual(
    name= "Condition of track",
                                                          breaks = c(0,1),
                                                          values = c("orange","steelblue2"),
                                                          labels= c("Normal","Wet"),
                                                          )+ylab("Proportion of unfinished drivers")+
  xlab("Rating of race")+geom_smooth(method = "lm", color = "red", se=FALSE)
```


We perform a t-test to check if there is enough statistical significance to 
support the hypothesis that the fan ratings do not differ for wet weather 
and normal weather races.

```{r}
t.test(datax[datax$wetweather == 0,]$RATING,datax[datax$wetweather == 1,]$RATING,alternative = 'l')
```

Based on the results we can see that the null hypothesis which states that there 
no difference in fan ratings for wet weather and normal weather races can be 
rejected at a 0.05 significance level.


```{r}
datax$wetweather <- as.factor(datax$wetweather)
mu <- datax %>% group_by(wetweather) %>% summarise_at(vars(RATING),list(name=mean))
mu
```
```{r}
image = jpeg::readJPEG('rain2.jpg')
ggplot(data=datax,aes(x=RATING,fill=factor(wetweather)))+theme_bw()+
  annotation_custom(rasterGrob(image = image,width = unit(1,"npc"), 
                               height = unit(1,"npc")), 
                               -Inf, Inf, -Inf, Inf)+geom_density(alpha=0.7)+
  scale_fill_manual(
    name= "Condition of track",
    values = c("orange","steelblue4"),labels=c("Normal","Wet")
  )+xlab("Rating of race")+
  ylab("Density")+
  geom_vline(data=mu,aes(xintercept=name,color = wetweather),alpha=2,linetype="twodash",size=1.3,show.legend=FALSE)+
  annotate(geom="text",label="p-value=<0.001",x=5,y=0.57)
```

```{r}
image = jpeg::readJPEG('rain2.jpg')
ggplot(data = datax,aes(y=RATING,x= wetweather))+
  annotation_custom(rasterGrob(image = image,width = unit(1,"npc"), 
                               height = unit(1,"npc")), 
                               -Inf, Inf, -Inf, Inf)+theme_bw()+
  geom_boxplot(aes(group=wetweather,fill=factor(wetweather)),alpha=1.5)+scale_fill_manual(
    name= "Condition of track",
    values = c("orange","steelblue2"),labels=c("Normal","Wet")
  )+xlab("Condition of Track")+
  ylab("Rating of race")+
  annotate(geom="text",label="p-value=<0.01",x=1.3,y=9.3)
```


```{r}
image = jpeg::readJPEG('rain3.jpg')
ggplot(data = datax,aes(sample=RATING,color=factor(wetweather)))+annotation_custom(rasterGrob(image = image,width = unit(1,"npc"), 
                               height = unit(1,"npc")), 
                               -Inf, Inf, -Inf, Inf)+theme_bw()+
  stat_qq()+scale_color_manual(breaks = c(0,1),name= "Condition of track",
                                                          values = c("orange","steelblue1"),
                                                          labels= c("Normal","Wet"),
                                                          )+
  geom_hline(yintercept = mean(datax[datax$wetweather==0,]$RATING),color="orange")+
  geom_hline(yintercept = mean(datax[datax$wetweather==1,]$RATING),color="steelblue2")+
  ylab("Quantiles of race rating")+xlab("Quantiles of normal distribution")

```
```{r}
dataxx <- datax %>% mutate(novertakes= (novertakes-mean(novertakes))/sd(novertakes))%>% 
  mutate(RATING = (RATING-mean(RATING))/sd(RATING)) %>% mutate(unfin = (unfin-mean(unfin))/sd(unfin))
lin=lm(RATING ~ novertakes+unfin+wetweather,data = dataxx)
summary(lin)
```
```{r}

name = c("Intercept","Number of overtakes","Unfinished drivers","wet weather")
coef = lin$coefficients
conf1 = confint(lin)[,1]
conf2 = confint(lin)[,2]
df1 <- data.frame(name, coef, conf1, conf2)
df1
```
```{r}
ggplot(data=df1, aes(x=coef,y=name))+geom_errorbar(width = .1, aes(xmin= conf1, xmax=conf2))+
  geom_point(shape= 21, size =3, fill="blue")+
  geom_vline(xintercept = 0, linetype="dashed", color="red")+xlab("Coefficient of linear regression")+
  ylab("Regressors")
```
```{r}
ggplot(data=df1, aes(x=coef,y=reorder(name,coef)))+geom_bar(stat = 'identity',fill=c("red3",'green2','red2','green3'))+
  geom_point(shape= 21, size =3, fill="blue")+
  geom_vline(xintercept = 0, linetype="dashed", color="red")+xlab("Coefficient of linear regression")+
  ylab("Regressors")
```

