---
title: "Visualisation of factors on excitement index calculated using youtube comments of a race."
output:
  pdf_document: default
  html_document:
    df_print: paged
  word_document: default
---

```{r}
library(dplyr)
library(plyr)
library(ggplot2)
library(gridExtra)
```

---
In this R program we use the dataset which contains data collected from the 
Ergast developer API for F1 races. and the excitement index calculated using the
positive and negative sentiments of the comments that are collected from youtube

The youtube comments are extracted using youtube Data V3 api.

The Stanford CoreNLP is used to get the sentiments for the extracted comments.

postiveSentiment in a comment is calculated as the ratio of number of positive 
sentences and total sentences in a comment.

negativeSentiment in a comment is calculated as the ratio of number of negative
sentiment sentences and total sentences in a comment.

All the positiveSentiments and negativeSentiments for all comments of a 
particular race are summed to give totalpos and totalneg for a particular race.

As not all extracted comments lead to a positive or negative sentiment we keep 
track of number of comments which result in a positive or negative sentiment as
number of useful comments.

And finally the "Excitement index" is calculated as the ratio between 
(totalpos+totalneg) and (number of useful comments).

We use this dataset to visualize the effects of 3 variables namely 
number of overtakes(novertakes), proportion of unfinished drivers
(unfin/totaldrivers) - the drivers who started the race but could not finish due
to various reasons and the weather of the race(wetweather) - F1 races are 
declared wet weather races if the weather conditions are wet/rainy/damp on the 
calculated "Excitement index".
---

```{r}
datax = read.csv("../Data/youtube_sentiments.csv")
colnames(datax)

```


```{r}
ggplot(data = datax,aes(x=Excitement_index,y= novertakes))+
  geom_point(aes(color=factor(wetweather)))+scale_color_manual(
    name= "Condition of track",
                                                          breaks = c(0,1),
                                                          values = c("orange","blue"),
                                                          labels= c("Normal","Wet"),
                                                          )+ylab("Number of overtakes in a race")+
  xlab("Excitement index of race")+
  geom_smooth(method = "lm", color = "red", se=FALSE)
```
```{r}
ggplot(data = datax,aes(x=Excitement_index,y= unfin))+
  geom_point(aes(color=factor(wetweather)))+scale_color_manual(
    name= "Condition of track",
                                                          breaks = c(0,1),
                                                          values = c("orange","blue"),
                                                          labels= c("Normal","Wet"),
                                                          )+ylab("NUmber of unfinished drivers")+
  xlab("Excitement index of race")+geom_smooth(method = "lm", color = "red", se=FALSE)
```


```{r}
ggplot(data = datax,aes(x=Excitement_index,y= unfin/totaldrivers))+
  geom_point(aes(color=factor(wetweather)))+scale_color_manual(
    name= "Condition of track",
                                                          breaks = c(0,1),
                                                          values = c("orange","blue"),
                                                          labels= c("Normal","Wet"),
                                                          )+ylab("Proportion of unfinished drivers")+
  xlab("Rating of race")+geom_smooth(method = "lm", color = "red", se=FALSE)
```


We perform a t-test to check if there is enough statistical significance to 
support the hypothesis that the fan ratings do not differ for wet weather 
and normal weather races.

```{r}
t.test(datax[datax$wetweather == 0,]$Excitement_index,datax[datax$wetweather == 1,]$Excitement_index,alternative = 'l')
```

Based on the results we can see that the null hypothesis which states that there 
no difference in fan ratings for wet weather and normal weather races can be 
rejected at a 0.05 significance level.


```{r}
datax$wetweather <- as.factor(datax$wetweather)
mu <- datax %>% group_by(wetweather) %>% summarise_at(vars(Excitement_index),list(name=mean))
mu
```
```{r}
ggplot(data=datax,aes(x=Excitement_index,fill=factor(wetweather)))+geom_density(alpha=0.5)+
  scale_fill_manual(
    name= "Condition of track",
    values = c("orange","blue"),labels=c("Normal","Wet")
  )+xlab("Excitement index of race")+
  ylab("Density")+
  geom_vline(data=mu,aes(xintercept=name,color = wetweather),linetype="dashed",show.legend=FALSE)
  #annotate(geom="text",label="p-value=",x=5,y=0.57)
```

```{r}
ggplot(data = datax,aes(y=Excitement_index,x= wetweather))+
  geom_boxplot(aes(group=wetweather,fill=factor(wetweather)))+scale_fill_manual(
    name= "Condition of track",
    values = c("orange","blue"),labels=c("Normal","Wet")
  )+xlab("Condition of Track")+
  ylab("Excitement index of race")
  #annotate(geom="text",label="p-value=0.0000001196",x=2.2,y=9.9)
```


```{r}
ggplot(data = datax,aes(sample=Excitement_index,color=factor(wetweather)))+
  stat_qq()+scale_color_manual(breaks = c(0,1),name= "Condition of track",
                                                          values = c("orange","blue"),
                                                          labels= c("Normal","Wet"),
                                                          )+
  geom_hline(yintercept = mean(datax[datax$wetweather==0,]$Excitement_index),color="orange")+
  geom_hline(yintercept = mean(datax[datax$wetweather==1,]$Excitement_index),color="blue")+
  ylab("Quantiles of race Excitement index")+xlab("Quantiles of normal distribution")

```
```{r}
dataxx <- datax %>% mutate(novertakes= (novertakes-mean(novertakes))/sd(novertakes))%>% 
  mutate(Excitement_index = (Excitement_index-mean(Excitement_index))/sd(Excitement_index)) %>% mutate(unfin = (unfin-mean(unfin))/sd(unfin))
lin=lm(Excitement_index ~ novertakes+unfin+wetweather,data = dataxx)
summary(lin)
```
```{r}

name = c("Intercept","Number of overtakes","Unfinished drivers","wet weather")
coef = lin$coefficients
conf1 = confint(lin)[,1]
conf2 = confint(lin)[,2]
df1 <- data.frame(name, coef, conf1, conf2)
df1
```
```{r}
ggplot(data=df1, aes(x=coef,y=name))+geom_errorbar(width = .1, aes(xmin= conf1, xmax=conf2))+
  geom_point(shape= 21, size =3, fill="blue")+
  geom_vline(xintercept = 0, linetype="dashed", color="red")+xlab("Coefficient of linear regression")+
  ylab("Regressors")
```

```{r}

# Create a dataframe only including the wet weather races
wet_weather_races <- datax[datax$wetweather==1,]

# Average total pos neg comments for wet weather races
avg_pos_neg_comments_wet <- (sum(wet_weather_races$totalpos)+
                              sum(wet_weather_races$totalneg))/
  sum(wet_weather_races$numberofusefulcomments)

# Average total pos neg comments for normal races
avg_pos_neg_comments_normal <-
  (sum(datax[datax$wetweather==0,]$totalpos)+
     sum(datax[datax$wetweather==0,]$totalneg))/
  sum(datax[datax$wetweather==0,]$numberofusefulcomments)

# Average viewcount for wet weather races
avg_viewcount_wet <- sum(wet_weather_races$viewcount)/dim(wet_weather_races)[1]

# Average viewcount for normal races
avg_viewcount_normal <- 
  sum(datax[datax$wetweather==0,]$viewcount)/
  dim(datax[datax$wetweather==0,])[1]


# Average likecount wet weather races
avg_likecount_wet <- sum(datax$likecount)/dim(datax)[1]

# Average likecount normal races
avg_likecount_normal <- 
  sum(datax[datax$wetweather==0,]$likecount)/
  dim(datax[datax$wetweather==0,])[1]

# Average dislikecount wet weather races
avg_dislikecount_wet <- sum(datax$dislikecount)/dim(datax)[1]

# Average dislikecount normal races
avg_dislikecount_normal <- 
  sum(datax[datax$wetweather==0,]$dislikecount)/
  dim(datax[datax$wetweather==0,])[1]

# Creating a dataframe with all the KPIs, so that they can be displayed in a
# plot.
comparison_wet_normal <- data.frame(c("Nass", "Normal", 
                                      "Nass", "Normal",
                                      "Nass", "Normal", 
                                      "Nass", "Normal"))

names(comparison_wet_normal) <- "race_type"

comparison_wet_normal$values <- c(avg_pos_neg_comments_wet, 
                                  avg_pos_neg_comments_normal, 
                                  avg_viewcount_wet, 
                                  avg_viewcount_normal,
                                  avg_likecount_wet, 
                                  avg_likecount_normal, 
                                  avg_dislikecount_wet,
                                  avg_dislikecount_normal)

comparison_wet_normal$column <- c("pos_neg_comments", "pos_neg_comments", 
                                  "viewcount", "viewcount",
                                  "likecount", "likecount",
                                  "dislike_count", "dislike_count")


# Create a plot for the viewcount
plot1 <- ggplot(data=comparison_wet_normal[comparison_wet_normal$column=="viewcount",],
                aes(x=column, y=values, fill=race_type)) + 
  geom_bar(stat="identity", position=position_dodge()) + scale_fill_manual(values=c("#B5074A", "#432D57")) +
  theme(legend.position = "none", axis.text.x = element_blank(),
        axis.ticks.x = element_blank()) + xlab("Durchschnittliche Zuschaueranzahl") + ylab("")

# Create a plot for likecount
plot2 <- ggplot(data=comparison_wet_normal[comparison_wet_normal$column=="likecount",],
                aes(x=column, y=values, fill=race_type)) + 
  geom_bar(stat="identity", position=position_dodge()) + scale_fill_manual(values=c("#B5074A", "#432D57")) +
  theme(legend.position = "none", axis.text.x = element_blank(),
        axis.ticks.x = element_blank()) + xlab("Durchschnittliche Anzahl an Likes") + ylab("")

# Create a plot for dislike count
plot3 <- ggplot(data=comparison_wet_normal[comparison_wet_normal$column=="dislike_count",],
                aes(x=column, y=values, fill=race_type)) + 
  geom_bar(stat="identity", position=position_dodge()) + scale_fill_manual(values=c("#B5074A", "#432D57")) +
  theme(legend.position = "none", axis.text.x = element_blank(),
        axis.ticks.x = element_blank()) + xlab("Durchschnittliche Anzahl an Dislikes") + ylab("")

# Create a plot for average number of positive/negative comments
plot4 <- ggplot(data=comparison_wet_normal[comparison_wet_normal$column=="pos_neg_comments",], 
                aes(x=column, y=values, fill=race_type)) + 
  geom_bar(stat="identity", position=position_dodge()) + scale_fill_manual(values=c("#B5074A", "#432D57")) +
  theme(legend.position = "none", axis.text.x = element_blank(), axis.ticks.x = element_blank())  +
  xlab("Anteil pos./neg. Kommentare") + ylab("")

# Create a legend for the plot
plot_legend <- ggplot(data=comparison_wet_normal[comparison_wet_normal$column=="pos_neg_comments",] 
                      ,aes(x=column, y=values, fill=race_type)) + 
  geom_bar(stat="identity",                                                 
           position=position_dodge()) + scale_fill_manual(values=c("#B5074A", "#432D57")) +
  theme(legend.position = "right", legend.direction = "horizontal", legend.title = element_blank())

# Some magic for creating a single legend for all the plots together
extract_legend <- function(my_ggp) {
  step1 <- ggplot_gtable(ggplot_build(my_ggp))
  step2 <- which(sapply(step1$grobs, function(x) x$name) == "guide-box")
  step3 <- step1$grobs[[step2]]
  return(step3)
}
shared_legend <- extract_legend(plot_legend)

# Create a grid with all the plots together and add the legend created above
grid.arrange(arrangeGrob(plot1,plot2,plot3,plot4, ncol = 2),
             shared_legend, nrow = 2, heights = c(10, 1))

```
